---
title: "World Happiness"
author: "Group 16"
date: "2021/6/30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

This contains all the libraries we need to load to run our code

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(infer)
library(GGally)
library(plotly)
library(RColorBrewer)
```

Read the data in to R and make sure it is a tibble.

```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
#Read "World Happiness" dataset into R.
setwd(dir = "D:\\STUDY\\GLASGOW\\Semester-2\\Data Analysis Skills\\Gruopwork\\Groupwork - 1 poster")
happy <- read_csv("happiness.csv")
happy <- as_tibble(happy)
```



# Introduction {#sec:Intro}

The World Happiness Dataset is the iconic statistical data of global happiness. It contains six critical factors: economic production, social support, life expectancy, freedom, absence of corruption, and generosity. To find which factors are most influential to the ladder score, we fitted a linear regression model with all relevant explanatory variables. Noting that Africa’s ladder score seems the smallest amongst all countries, we used bootstrapping to find the 95th percentile of the difference in ladder score between Africa and the rest of the world and fit the regression model for these two cases.

## Explanation of the Variables

Outcome Variable

Ladder Score: the scale of happiness, which is called happiness score in the poster


Explanatory Variables

Logged GDP per capita: the scale of economic production

Social support: the scale of social support

Healthy life expectancy: the scale of life expectancy

Freedom to make life choices: the scale of freedom

Generosity: the scale of generosity

Perceptions of corruption: the scale of absence of corruption

# Explanatory analysis {#sec:Explan}

This code adds a column to the data, using the mutate function, so each country is either Africa or Rest of the World (catergory).

```{r addcol}
#Add a column called African, if countries are in Africa, the value is Africa, otherwise, the value is World.
happy <- happy %>%
  mutate(African = ifelse(happy$`Regional indicator` == "Sub-Saharan Africa" | happy$`Regional indicator` == "Middle East and North Africa",
                "Africa","World"))

```


This code shows a summary for each region. The mean, SD, quantiles, min, max and median values are all calculated.
The statistics relate to the ladder score, with is the score for happiness. (higher being better score).

North America and ANZ's mean happiness score is 7.13 which is the highest score. Western Europe's mean score is 6.91 which is the second highest score. South Asia's mean score is at the bpttom of the table at 4.44. Sub-Saharan Africa is the second lowest on the list at 4.49.


```{r summ}
#Give summary to Happiness score, including n, mean, standard deviation, 1st quantile, 3rd quantile, min and max value, median value.
happy %>%
  group_by(`Regional indicator`) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25), TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`)) %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```


The graph shows the sactterplot of Happniess score in each region. There is a great difference in the distribution of Western Europe and Sub-Saharan Africa.
```{r}
ggplot(data = happy,mapping = aes(y=`Regional indicator`,x=`Ladder score`,color=`Regional indicator`))+
  geom_point()+
  labs(x="Happiness score",y="Regions")+
  theme(legend.position="none")
```

Also summary data for Africa and Rest of the World.
African countries have 53 and 96 others. African countries scored 4.73 in average, much lower than other countries'  mean scores of 5.98.
```{r sumafrica}
#summary stats for AFRICA and REST OF WORLD
happy %>%
  group_by(African) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25),
            TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`))  %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")


```


This code uses bootstrapping to find a 95% confidence interval for the mean ladder score of all the data. By setting the seed to 1, it ensures that the bootstrap gives the same distribution each time, and hence it is reproducible. 

The bule region shows the  95% confidence interval for mean happiness score distribution. It is between 5.36 and 5.70.

```{r bootstrap}
set.seed(1)
bootstrap_distribution_mean <- happy %>% 
  specify(response = `Ladder score`) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean")

bootstrap_distribution_mean %>% 
  get_ci(level = 0.95, type = "percentile")

percentile_ci <- bootstrap_distribution_mean %>% 
  get_ci()

boots<-bootstrap_distribution_mean %>% 
  visualize(endpoints = percentile_ci, direction = "between")
boots

```


This next section of code produces a boxplot of ladder score, using region as a factor for all regions.

```{r box, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:scat} Ladder score by region.", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(`Regional indicator`), 
                                         y = `Ladder score`)) + geom_boxplot(fill = "slateblue") + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = unique(happy$`Regional indicator`))+
  theme(axis.text.x = element_text(angle=45,hjust = 0.5,vjust = 0.5))

```


Now same boxplot but just for Africa and rest of the World

```{r boxaf, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:bplot} Ladder score by Africa", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(African), 
                                   y = `Ladder score`,fill = "transparent"))+ 
  geom_boxplot(fill = brewer.pal(7, "Set1")[3]) + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = c("Africa","Rest of World"))

```


Using bootstrapping (week 7), we can find a 95th percentile of the difference in ladder score between Africa and rest of World.  
This shows that the score is between 1.65 and 1.11 lower in Africa than rest of World.

```{r boot}

bootstrap_distribution <- happy %>%
  specify(`Ladder score`~ African) %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in means", order = c("Africa", "World"))
percentile_ci <- bootstrap_distribution %>%
  get_ci()

percentile_ci

```

# Linear models {#sec:models}

## Question 1 
We have six variables and want to know is there statistically significant relationship between happiness score and six variables and use these to estimate the model.

First we fit the full model includes six variables.

$$\widehat{\mbox{Ladder score}} = \widehat\alpha + \widehat\beta_{\mbox{1}} \cdot \mbox{GDP}
+\widehat\beta_{\mbox{2}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{3}} \cdot \mbox{life expectancy} + \widehat\beta_{\mbox{4}} \cdot \mbox{Freedom} + \widehat\beta_{\mbox{5}} \cdot \mbox{Generosity} + \widehat\beta_{\mbox{6}} \cdot \mbox{corruption}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,...,6);

```{r full_model}
full_model<-lm(`Ladder score`~`Logged GDP per capita`+`Social support`+`Healthy life expectancy`+`Freedom to make life choices`+`Generosity`+`Perceptions of corruption`,data = happy)
get_regression_table(full_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')

summary(full_model)
```

It shows that variables "Logged GDP per capita", "Social support", and "Freedom to make life choice"'s have lowest p_value which are all less than 0.05 so they are most significant. We also check the correlationship between these six variables.

```{r correlation}
ggpairs(happy[,c(3,7:12)])
```

Although Healthy life expectancy" has the strongest correlation with "Ladder score", this variable also has the high correlation with other two variables (Logged GDP per capita and Social support).

Finally, we decide to use three variables, Logged GDP per capita, Social support and Freedom to make life choices to fit our model.

$$\widehat{\mbox{Happiness score}} = \widehat\alpha + \widehat\beta_{\mbox{1}} \cdot \mbox{GDP}
+\widehat\beta_{\mbox{2}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{3}} \cdot \mbox{Freedom}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,2);

```{r final_model}
final_model<-lm(`Ladder score`~`Social support`+`Freedom to make life choices`+`Logged GDP per capita`,data = happy)
summary(final_model)
get_regression_table(final_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the final fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')
summary(final_model)
```

Our final model shows as follows,

$$\widehat{\mbox{Happiness score}} = -2.61+ 0.44 \cdot \mbox{GDP}
+2.34 \cdot \mbox{Social support} + 2.68 \cdot \mbox{Freedom}, $$

Before we can proceed to use the fitted model, we must check the assumption of the model. We assesses the model assumptions by plotting the residuals against the explanatory variable in Figure \ref{fig:resids_1} ,\ref{fig:resids_1} and a histogram (Figure \ref{fig:hist}) of the residuals to assess whether they are normally distributed.

```{r assumptions}
regression_points <- get_regression_points(final_model)
```

Check the first assumption of variable "Social support".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_1} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Social support`,y=residual)) +
  geom_point() +
  geom_jitter(width = 0.1) +
  labs(x = "Social support",y="Residual")+
  geom_hline(yintercept = 0,col="blue",size=1)
```

Check the first assumption of variable "Freedom to make life choices".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_2} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Freedom to make life choices`, y = residual)) +
  geom_point() +
  labs(x = "Freedom to make life choices", y = "Residual", title = "Residuals vs Freedom to make life choices")  +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```

Check the first assumption of variable "Logged GDP per capita".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_2} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Logged GDP per capita`, y = residual)) +
  geom_point() +
  labs(x = "Logged GDP per capita", y = "Residual", title = "Residuals vs Freedom to make life choices")  +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```

The scatterplots show an approximately spread of the residuals above and below the zero line, which means the data statisfied the first assumption, the residuals have mean zero.

```{r hist,echo=FALSE,fig.with=13,fig.align="center",fig.cap="\\label{fig:hist} Histgram of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points,aes(x=residual))+
  geom_histogram(color="white")+
  labs(x="Residuals")
```

The histgram support the assumption of normally distributed errors in the model.

## Question 2

It is interesting to look at how corruption affects the happiness score for Africa compared to the rest of the World, and see if there are any differences.

Fitting an interaction model with separate intercepts and gradients.

$$\widehat{\mbox{Happiness score}} = \widehat\alpha 
+\widehat\beta_{\mbox{1}} \cdot \mbox{Region} + \widehat\beta_{\mbox{2}} \cdot \mbox{Corruption}+\widehat\beta_{\mbox{3}} \cdot \mbox{Region}\cdot \mbox{Corruption}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,2,3);

```{r}
#fit an interaction model
Afr.model <- lm(`Ladder score` ~ `Perceptions of corruption` * African, data = happy)
summary(Afr.model)
get_regression_table(Afr.model)
```

Our final estimated interaction model shows as follows,

For African countries:

$$\widehat{\mbox{Happiness score}} = 4.682+ 0.06 \cdot \mbox{Corruption} $$

For Non-African countries:

$$\widehat{\mbox{Happiness score}} = 7.68 - 2.462 \cdot \mbox{Corruption} $$

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:corruption} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
#Draw the graph to show that corruption variable have different impact on Afircan countries and Non-African countries.
ggplot(happy, aes(x = `Perceptions of corruption`, y = `Ladder score`, color = African)) +
  geom_jitter() +
  labs(x = "Perceptions of corruption", y = "Ladder Score", color = "African") +
  geom_smooth(method = "lm", se = FALSE)
```
The above figure shows that for the rest of the World, increased corruption is associated with lower ladder scores. However, in Africa, there is little evidence that an increase in corruption makes much difference to happiness.

This suggests that for the rest of the World, corruption is an important issue that can significantly reduce happiness. In Africa, this variable is less significant, perhaps due to other more fundamental problems like life expectancy or GDP being more important.

Before we can proceed to use the fitted model, we must check the assumption of the model.

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:assumption} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
regression.points <- get_regression_points(Afr.model)
regression.points

ggplot(regression.points, aes(x = `Ladder score`, y = residual)) +
  geom_point() +
  labs(x = "Ladder score", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~ African)
  
ggplot(regression.points, aes(x = residual)) +
  geom_histogram(binwidth = 0.25, color = "white") +
  labs(x = "Residual") +
  facet_wrap(~African)

```

# Conclusions {#sec:Conc}
In question 1, by fitting the full model for six variables (critical factors) and comparing correlation and p-value with "Happiness score", we found the "Logged GDP per capita", "Freedom to make life choice" and "Social support" are most significant and also have the relatively high correlations with response variable “Happiness score". Hence, we chose those three variables (the residuals all have mean zero), "Logged GDP per capita", "Social support" and "Freedom to make life choices" to fit our model. There are moderately strong positive linear relationships between "Happiness score" and "Social support". Furthermore, the linear relationships between "Happiness score" and "Social support" in the most non-African countries is stronger than that in African countries. Similarly, non-African countries have higher "Freedom to make life choice" and "Logged GDP per capita" compared with African countries.

In question 2, There is a noticeable difference in the gradient for the rest the of the World than for Africa. For the rest of the World, increased "Corruption" is associated with lower "Happiness score". However, in Africa, there is little evidence that an increase in corruption makes much difference to happiness, which means "Corruption" has less impact on African countries, while for the rest of the world, this variables will significantly reduce happiness.
